Predictions From Day 0 Community Structures

### 2016-07-18 code/HumanCdGF_microbiota_predict.Rmd
Split data into test/train for randomforest

#create function for random forest
set.seed(1)
run_rf_on <- function(DATAFRAME, RESPONSE, OTUs = OTU_list){
     Predictor_df <- as.data.frame(DATAFRAME[!DATAFRAME[,RESPONSE] %in% NA,c(RESPONSE,OTUs)])
     RF <- randomForest(as.formula(paste(RESPONSE, '~ .')), data = Predictor_df, importance = TRUE)
     R2 <- rSquared(Predictor_df[,RESPONSE], Predictor_df[,RESPONSE] - predict(RF,Predictor_df))
     MSE <- mean((Predictor_df[,RESPONSE] - predict(RF,Predictor_df))^2)
     IMPORT <- importance(RF)
     FOREST <- RF
     PLOT<- ggplot(data = data.frame(actual=Predictor_df[,RESPONSE],pred=predict(RF,Predictor_df)),
       aes(x=actual, y=pred)) + geom_point() + geom_abline(color='red')
     return(list(rf=FOREST,r2=R2,mse=MSE,import=IMPORT,plot=PLOT))
}
```

##Can we predict the endpoint C. difficile CFU with the community structure at Day 0?
But endpoint has data from Day 2,3,10
     Might be misleading since contains CFU from different data points

@@

% Variance explained `r Day_Euth_cdiff_cfu_rf$rf$rsq[length(Day_Euth_cdiff_cfu_rf$rf$rsq)]*100`

@@

##Can we predict moribundity with the community structure at Day 0?
Need to convert day euthanized to binary categorical
```{r, echo=FALSE}
##Fix to be categorical, and not regression
Predict_early_euth_df <- Predict_df[!Predict_df$Euth_Early %in% NA,c('Euth_Early',OTU_list)]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df)
rf_early_euth
```

@@

```{r, echo=FALSE}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot

@@

There seems a flucuation in the % variance explained, which seems to coincide with the flutuations in cdiff cfu

@@

>>>>

#create function for random forest
run_rf_on <- function(DATAFRAME, RESPONSE, SAMPLE_SPLIT = 2/3,  OTUs = OTU_list){
set.seed(2)
     #
     Predictor_df <- as.data.frame(DATAFRAME[!DATAFRAME[,RESPONSE] %in% NA,c(RESPONSE,OTUs)])
     sample_set <- sample(seq_len(nrow(Predictor_df)), size = floor((SAMPLE_SPLIT)*nrow(Predictor_df)))
     test <- Predictor_df[-sample_set,]
     train <- Predictor_df[sample_set,] 
          RF <- randomForest(as.formula(paste(RESPONSE, '~ .')), data = train, importance = TRUE)
     R2 <- rSquared(test[,RESPONSE], test[,RESPONSE] - predict(RF,test))
     MSE <- mean((test[,RESPONSE] - predict(RF,test))^2)
     IMPORT <- importance(RF)
     FOREST <- RF
     PLOT<- ggplot(data = data.frame(actual=test[,RESPONSE],pred=predict(RF,test)),
       aes(x=actual, y=pred)) + geom_point() + geom_abline(color='red')
     ##MEAN_DECREASE <- varImpPlot(RF)
     return(list(rf=FOREST,r2=R2,mse=MSE,import=IMPORT,plot=PLOT))
}
```

##Can we predict the endpoint C. difficile CFU with the community structure at Day 0?
But endpoint has data from Day 2,3,10

Might be misleading since contains CFU from different data points

@@

% Variance explained `r Day_Euth_cdiff_cfu_rf$rf$rsq[length(Day_Euth_cdiff_cfu_rf$rf$rsq)]*100`  
r^2^ = `r Day_Euth_cdiff_cfu_rf$r2`  
MSE = `r Day_Euth_cdiff_cfu_rf$mse`

@@

##Can we predict moribundity with the community structure at Day 0?
```{r, echo=FALSE}
##Fix to be categorical, and not regression
Predict_early_euth_df <- Predict_df[!Predict_df$Euth_Early %in% NA,c('Euth_Early',OTU_list)]
rf_early_euth <- randomForest(as.factor(Euth_Early) ~ ., data = Predict_early_euth_df)
rf_early_euth
```

@@

```{r, echo=FALSE}
Day_10_cdiff_cfu_rf <- run_rf_on(Predict_df,'Day_10_cdiff_cfu')
Day_10_cdiff_cfu_rf$plot
```

@@

There seems a flucuation in the % variance explained, which seems to coincide with the fluctuations in cdiff cfu

##seed greatly affects variance, likely due to data split, sample size might be too small to make reliable predictions

@@

***
***
